Checking Big Suffix and LCP Arrays by Probabilistic Methods

### 2017-03-14 draft/draft.tex
2017/03/14

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as the fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. In this paper, we propose two methods to check the suffix and LCP arrays in external memory by using a Karp-Rabin fingerprinting technique, where the checking result is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two neighboring suffixes in the given suffix array by computing and comparing the fingerprints of their LCPs. This idea is also employed in the second method to verify a subset of the given suffix and LCP arrays, from which then a copy of the final suffix and LCP arrays is produced following the induced sorting principle and compared with the given arrays for verification.

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
	Suffix and LCP arrays verification, Karp-Rabin fingerprinting technique, external memory.
\end{IEEEkeywords}}

@@

Suffix and longest common prefix (LCP) arrays play an important role in various string processing tasks, such as data compression, pattern matching and genome assembly. In many applications, these two data structures make up the core part of a powerful full-text index, called enhanced suffix array~\cite{Abouelhodaa2004}, which is more space efficient than a suffix tree and applicable to emulating any searching functionalities provided by the latter in the same time complexity. The first algorithm for building suffix array in internal memory was presented in~\cite{Manber1993}. From then on, much more effort has been put on designing efficient constructors for suffix array~(SA) on different memory models~\cite{Karkkainen2003, Ko2003, Kim2003, Nong11, Dementiev2008, Ferragina2012, Manzini2004, Bingmann12, Karkkainen2014, Nong14, Nong15}. In respect of the research on LCP array construction algorithms, the existing works can be classified into two categories with regard to their input requirements, where the algorithms from the first category compute both suffix and LCP arrays at the same time with the original text only~\cite{Fischer11, Bingmann12, Flick2015} and those from the second category carry out the computation by taking SA and/or Burrows-Wheeler transform (BWT) as additional input~\cite{Kasai2001,Karkkainen2009, Fischer11, Puglisi2008, Deo2013}. Among all, the algorithms designed by the induced sorting~(IS) principle take linear time and space to run and outperform previous arts on both internal and external memory models~\cite{Nong11, Karkkainen2014}. Recently, there appear some novel works that are competitive with the IS-based ones and can even achieve better performance when adequate computation resources are at hand. These algorithms are specific for parallel computation environments and capable of achieving high performance by fully using the available multi-core CPUs and/or GPUs~\cite{Osipov2012, Deo2013, Wang2015, Karkkainen2015, Karkkainen2016}.

While the research on efficient construction of suffix and LCP arrays is evolving, the algorithms proposed recently are becoming more complicated than before. This reveals a need for program debugging because a program gives no guarantee that it has correctly implemented the underlying algorithm. As a common practice, a suffix or LCP checker is provided to check the correctness of a constructed array. For example, such a checker can be found in some software packages for SA-IS~\cite{Nong11}, eSAIS~\cite{Bingmann12}, DC3~\cite{Dementiev08} and so forth. In addition to help avoid implementation bugs, a checker is also demanded for an array constructed by a probabilistic algorithm~(e.g.~\cite{Bille2013}). In this case, the array is correctly constructed with a probability and hence must be verified by a checker to ensure its correctness. As far as we know, the work in~\cite{Burkhardt2003} describes the only SA checking method that can be found in the existing literature, and no efficient checking method for LCP array has been reported yet. Particularly, there is currently no reported solution that can check both the suffix and the LCP arrays in external memory. This motivates our work here to design high-performance checkers for massive data.

@@

Our contribution mainly includes two methods to probabilistically verify the given suffix and LCP arrays. Method A checks the lexical order and the LCP-value of two neighboring suffixes in SA by literally comparing the characters of their LCPs in pairs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check equality of these sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. When implemented in external memory, it suffers from a space bottleneck owing to the large disk volume taken by each sort. 
To overcome the drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, then it reuses the inducing process of an IS-based construction algorithm to produce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experiments indicate that the peak disk use of the program for the algorithm designed by Method B is only half as that of the program for the algorithm designed by Method A.

The remainder of this paper is organized as follows. We first describe the two methods and their algorithmic designs in Sections~\ref{sec:method1} and~\ref{sec:method2}, and then the evaluation of our programs for the algorithms designed by these methods in Section~\ref{sec:experiment}. Finally, we present the concluding remarks in Section~\ref{sec:conclusion}.

@@

Given a string $x[0, n)$ drawn from an alphabet $\Sigma$, the suffix array of $x$, denoted by $sa$, is a permutation of $\{0, 1, ..., n - 1\}$ such that ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ for $i, j \in [0, n)$ and $i < j$, where ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$ are two suffixes starting with $x[sa[i]]$ and $x[sa[j]]$, respectively. Particularly, we say ${\sf suf}(sa[j])$ is a lexical neighbor of ${\sf suf}(sa[i])$ if $|i - j| = 1$. The LCP array of $x$, denoted by $lcp$, consists of $n$ integers, where $lcp[0]=0$ and $lcp[i]$ records the LCP-value of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$ for $i \in [1, n)$.

@@

According to the above definitions, we show in Lemma~\ref{lemma:1} the sufficient and necessary conditions for checking suffix and LCP arrays. Notice that the lexical order and the LCP-value of any two suffixes in $x$ can be computed by literally comparing their characters from left to right. Because all the suffixes differ in length and end with a common character, there must exist $k \in [0, n)$ such that $x[i, i + k) = x[j, j + k)$ and $x[i + k] \ne x[j + k]$ for any $i, j \in [0, n)$ and $i \ne j$. This approach can be also applied to verifying the last two conditions in Lemma~\ref{lemma:1}, but it takes at worst $\mathcal{O}(n)$ character-wise comparisons for each pair of neighboring suffixes in $sa$ to compare the two underlying substrings indicated by their LCP-value.

@@

An alternative is to exploit a perfect hash function~(PHF) to convert each substring into a single integer such that any two substrings have a common hash value if and only if they are literally equal to each other. This implies that we can compare the hash values instead to check equality of their corresponding substrings. The key point here is how to quickly calculate the hash values of $x[sa[i], sa[i] + lcp[i] - 1]$ and $x[sa[i - 1], sa[i - 1] + lcp[i] - 1]$ for all $i \in [1, n)$. Taking into consideration the high difficulty of finding a PHF to meet this requirement, we prefer using a Karp-Rabin fingerprinting function~\cite{Karp1987} to transform a substring into its integer form, called fingerprint. To be specific, suppose $L$ is a prime and $\delta$ is a number randomly chosen from $[1, L)$, the fingerprint ${\sf fp}(i, j)$ for a substring $x[i, j]$ can be calculated according to the formulas below as following: scan $x$ rightward to iteratively compute ${\sf fp}(0, k)$ for all $k \in [0, n)$ using Formulas~\ref{formula:1}-\ref{formula:2}, record ${\sf fp}(0, i - 1)$ and ${\sf fp}(0, j)$ during the calculation and subtract the former from the latter to obtain ${\sf fp}(i, j)$ using Formula~\ref{formula:3}.

@@

We point out that two equal substrings always share a common fingerprint, but the inverse is not true. Fortunately, it has been proved in~\cite{Karp1987} that the probability of a false match can be reduced to a negligible level by setting $L$ to a large value\footnote{This property is utilized in~\cite{Bille2013} to design a probabilistic algorithm for computing a sparse suffix array. }. This leads us to the conclusion in Corollary~\ref{corollary:1}.

@@

We close this part with an example in Fig.~\ref{fig:example} for a better understanding of computing and comparing the fingerprints of the substrings specified the LCP-values of lexical neighbors. Given that $L = 197$ and $\delta = 101$, lines 4-8 compute ${\sf fp}(0, p)$ iteratively according to Formulas~\ref{formula:1}-\ref{formula:2}. Then, lines 10-16 use these values to compute the fingerprints for target substrings. Consider the leftmost pair of neighboring suffixes in $sa$, that is ${\sf suf}(sa[0])$ and ${\sf suf}(sa[1])$, the substrings indicated by their LCP-value are $x[sa[0], sa[0] + lcp[1] - 1]$ and $x[sa[1], sa[1] + lcp[1] - 1]$, respectively. According to Formula~\ref{formula:3}, ${\sf fp}(sa[0], sa[0] + lcp[1] - 1)$ is equal to the difference between ${\sf fp}(0, sa[0] - 1)$ and ${\sf fp}(0, sa[0] + lcp[1] - 1)$, both of which have been calculated beforehand. Hence, we obtain the fingerprints for these two substrings in lines 10-12 and see that they are equal to each other.

@@

	\includegraphics[width = 0.9\columnwidth]{example}
	\caption{An Example for Computing and Comparing Fingerprints for Substrings Specifed by the LCP-Values of Neighboring Suffixes in SA. \label{fig:example}}	
\end{figure}

@@

It is clear that the above algorithm consumes $\mathcal{O}(n)$ time and space when implemented in internal memory. However, if the two auxiliary arrays cannot be wholly accommodated into RAM during the execution of S2, it suffers from a performance degradation caused by frequent random accesses to disks. Assume that $x$, $sa$ and $lcp$ are stored in external memory, we design Algorithm~\ref{alg:1} for conducting these I/O operations in a disk-friendly way. The main idea is to sort data in the order that they are visited and then access them by sequential reads. For the purpose, Algorithm~\ref{alg:1} first scans $sa$ and $lcp$ to produce $ST_1, ST_2, ST_3$ and sorts their tuples by 1st component in ascending order at the very beginning (lines~\ref{alg:1:a}-\ref{alg:1:b}). Afterward, it iteratively computes the fingerprints of all the prefixes according to Formulas~\ref{formula:1}-\ref{formula:2} and assigns them to the sorted tuples as following (lines~\ref{alg:1:c}-\ref{alg:1:d}): when figuring out ${\sf fp}(0, i - 1)$, extract each tuple $e$ with $e.1st = i$ from $ST_1/ST_2/ST_3$, update $e$  with ${\sf fp}(0, i - 1)$, and then forward $e$ to $ST_1'/ST_2'/ST_3'$. Because the 1st components of the tuples in $ST_1$ constitute a copy of $sa$, the algorithm can check condition (1) when scanning these tuples in their sorted order. Finally, it sorts the updated tuples back to their original order~(line~\ref{alg:1:e}) and visits them  sequentially to check conditions (2)-(3) following the same way of S2~(lines~\ref{alg:1:f}-\ref{alg:1:g}).

@@

			$fp := fp \cdot \delta + x[i] \! \mod \! P$
		} \label{alg:1:d}

@@

			\If{$\hat{fp_1} \ne \hat{fp_2}$ {\sf or} $ch_1 \le ch_2$}{

@@

The last point to be mentioned here is how to obtain the value of $\delta^{lcp[i]}$ quickly when computing $\hat{fp_1}$ and $\hat{fp_2}$ in lines~\ref{alg:1:j} and~\ref{alg:1:k}. One method is to keep a lookup table in internal memory to store $\delta^{lcp[i]}$ for all $i \in [0, n)$. This method can answer the question in constant time, but it is space-consuming and 
impractical to use if the table size exceeds the capacity of memory bank. We provide another method that returns the answer in $\mathcal{O}(\lceil \log2^n \rceil)$ time using $\mathcal{O}(\lceil \log2^n \rceil)$ internal memory space. Specifically, suppose $e$ is an integer from $[0, n)$, it can be broken down into the form of $\Sigma_{i = 0}^{\lceil \log2^n \rceil}{k_i \cdot 2^i}$ by performing at most $\lceil \log2^n \rceil$ divisions, where $k_0, k_1, ..., k_{\lceil \log2^n \rceil}$ are integers from $\{0, 1\}$ . Thereby, we have $\delta^e = \Pi_{i = 0}^{\lceil \log2^n \rceil}{\delta}^{k_i \cdot 2^i}$ by replacing $e$ with its decomposition, where the expression on the right side of the equation can be easily computed with $\{{\delta}^{1}, {\delta}^{2}, \dots, {\delta}^{2^{\lceil \log2^n \rceil}} \}$ already known.

@@

Algorithm~\ref{alg:1} is I/O-intensive, it performs multiple scans and sorts on the arrays of $\mathcal{O}(n)$ fixed-size tuples residing on disks. Given an external memory model with RAM size $M$, disk size $D$ and block size $B$, all are in words, the time and I/O complexities for each scan are $\mathcal{O}(n)$ and $\mathcal{O}(n / B)$, respectively, while those for each sort are $\mathcal{O}(n\log_{M/ B}(n / B))$ and $\mathcal{O}((n / B)\log_{M / B}(n / B))$, respectively~\cite{Arge2013}. Clearly, this algorithm reaches its peak disk use when sorting tuples in lines~\ref{alg:1:b} and~\ref{alg:1:e}. The experimental study in Section~\ref{sec:experiment} indicates that our program for Algorithm~\ref{alg:1} is rather space consuming, its peak disk use reaches 40 bytes per input character.

@@

In this part, we describe an alternative checking method based on the induced sorting principle. Compared with Algorithm~\ref{alg:1}, our program for the algorithm designed by this method only takes half space on real-world datasets. Before the presentation, we introduce some symbols and notations used in the following paragraphs.

@@

{\em Suffix and LCP buckets.} Suppose $sa$ is correct, all the suffixes in $sa$ are naturally partitioned into multiple buckets and those of a common heading character are grouped into a single bucket that occupies a contiguous interval in $sa$. Each bucket can be further divided into two sub-buckets, where the left and right parts only contain L- and S-type suffixes, respectively. For short, we use ${\sf sa\_bkt}(c)$ to denote the bucket storing suffixes starting with character $c$ and ${\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$ to denote its left/right sub-bucket. Accordingly, $lcp$ can be also split into multiple buckets, where ${\sf lcp\_bkt}(c)/{\sf lcp\_bkt_L}(c)/{\sf lcp\_bkt_S}(c)$ stores the LCP-values of suffixes in ${\sf sa\_bkt}(c)/{\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$.

{\em Suffix and LCP arrays for S*-type suffixes.} Given that the number of S*-type suffixes is $n_1$, $sa^*[0, n_1)$ stores all the S*-type suffixes and arrange them in lexical order, while the $(i + 1)$-th item of $lcp^*[0, n_1)$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$.

@@

The induced sorting principle has been extensively used to design efficient algorithms for constructing the suffix and LCP arrays on internal and external memory models. An IS-based construction algorithm mainly consists of a reduction phase for computing $sa^*$ and $lcp^*$, followed by an induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$~\footnote{In the interest of completeness, we give an overview of the induction phase in Appendix~\ref{sec:appendix}.}. Given that $sa^*$ and $lcp^*$ are already known, we can produce the final suffix and LCP arrays by calling the inducing process of an existing construction algorithm. This enlightens us to design a checker based on Lemma~\ref{lemma:2}.

@@

\begin{enumerate}[(1)]
	\item
	$sa^*$ and $lcp^*$ are both correct.
	\item
	$sa = sa'$ and $lcp = lcp'$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by calling the inducing process of any existing IS-based construction algorithm.
\end{enumerate}
\end{Lemma}

Similar to Method A, we can employ the fingerprinting technique to probabilistically check the correctness of $sa^*$ and $lcp^*$. Based on this idea, we describe in the next subsection an external-memory algorithm for checking the conditions in Corollary~\ref{corollary:2}.

\begin{Corollary} \label{corollary:2}
Both $sa[0, n)$ and $lcp[0, n)$ are correct with a high probability given the following conditions, for $i \in [0, n)$, $\{j, k\} \in [1, n_1)$ and $j \ne k$:

\begin{enumerate}[(1)]
	\item
	$sa^*[j] \ne sa^*[k]$.
	\item
	${\sf fp}(sa^*[j], sa^*[j] + lcp^*[j] - 1) = {\sf fp}(sa^*[j - 1], sa^*[j - 1] + lcp^*[j] - 1)$.
	\item
	$x[sa^*[j] + lcp^*[j]] > x[sa^*[j - 1] + lcp^*[j]]$.
	\item
	$sa[i] = sa'[i]$ and $lcp[i] = lcp'[i]$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by calling the inducing process of an existing IS-based construction algorithm.
\end{enumerate}

\end{Corollary}

@@

	sort tuples in $ST_2$ by 2nd component \label{alg:2:g}

@@

	\If{${\sf CheckByFP}(x, sa^*, lcp^*, lcp^*.{\sf size}()) = false$}{  \label{alg:2:i}

		\Return false;  \hspace{1cm} // conditions (2) or (3) is violated
	}\label{alg:2:j}

@@

			\Return false	// condition (4) is violated
		}	
	}\label{alg:2:m}

>>>>

For full-text indexing of massive data, the suffix and LCP (longest common prefix) arrays have been recognized as fundamental data structures, and there are at least two needs in practice for checking their correctness, i.e. program debugging and verifying the arrays constructed by probabilistic algorithms. In this paper, we propose two probabilistic methods to check the suffix and LCP arrays of constant or integer alphabets in external memory by using a Karp-Rabin fingerprinting technique, where the checking result is wrong only with a negligible error probability. The first method checks the lexicographical order and the LCP-value of two suffixes by computing and comparing the fingerprints of their LCPs. This method is rather general in terms of that it can verify any full or sparse suffix/LCP array of any order. The second method is more space efficient, it first employs the fingerprinting technique to verify a subset of the given suffix and LCP arrays, from which then a copy of the suffix and LCP arrays is produced using the induced sorting principle and compared with the given arrays for verification, where the copy of the induced suffix and LCP arrays can be removed for constant alphabets.

@@

% Note that keywords are not normally used for peerreview papers.
\begin{IEEEkeywords}
	Suffix and LCP arrays, verification, Karp-Rabin fingerprinting, external memory.
\end{IEEEkeywords}}

@@

Suffix and longest common prefix (LCP) arrays play an important role in various string processing tasks, such as data compression, pattern matching and genome assembly. In many applications, these two data structures make up the core part of a powerful full-text index, called enhanced suffix array~\cite{Abouelhodaa2004}, which is more space efficient than a suffix tree and applicable to emulating most searching functionalities provided by the latter in the same time complexity. The first algorithm for building suffix array~(SA) in internal memory was presented in~\cite{Manber1993}. From then on, much more effort has been put on designing efficient constructors for suffix array on different computation models~\cite{Karkkainen2003, Ko2003, Kim2003, Nong11, Dementiev2008, Ferragina2012, Manzini2004, Bingmann12, Karkkainen2014, Nong14, Nong15}. In respect of the research on LCP array construction algorithms, the existing works can be classified into two categories with regard to their input requirements, where the algorithms from the first category compute both suffix and LCP arrays at the same time with the original text only~\cite{Fischer11, Bingmann12, Flick2015}, and those from the second category carry out the computation by taking SA and/or Burrows-Wheeler transform (BWT) as additional inputs~\cite{Kasai2001,Karkkainen2009, Fischer11, Puglisi2008, Deo2013}. So far, the algorithms designed by the induced sorting~(IS) principle take linear time and space to run and get the best results on both internal and external memory~\cite{Nong11, Karkkainen2014, Karkkainen2017}. In addition to the sequential algorithms, there are also parallel algorithms proposed to achieve high performance by fully using the available multi-core CPUs and/or GPUs~\cite{Osipov2012, Deo2013, Wang2015, Karkkainen2015, Karkkainen2016}.

While the research on efficient construction of suffix and LCP arrays keeps evolving, the algorithms proposed recently are becoming more complicated than before. Currently, the open source programs for the state-of-the-art algorithms are provided "as-is" for demonstration and experiment purpose only, giving no guarantee that they have correctly implemented the algorithms. As a common practice, a suffix or LCP checker is provided to check the correctness of a constructed array. For example, such a checker can be found in some software packages for DC3~\cite{Dementiev08}, SA-IS~\cite{Nong11}, eSAIS~\cite{Bingmann12} and so forth. In addition to help avoid implementation bugs, a checker is also demanded for an array constructed by a probabilistic algorithm~(e.g.~\cite{Bille2013}). In this case, the array is correctly constructed with a probability and hence must be verified by a checker to ensure its correctness. As far as we know, the work in~\cite{Burkhardt2003} describes the only SA checking method that can be found in the existing literature, and no efficient checking method for LCP array has been reported yet. Particularly, there is currently no reported solution that can check both the suffix and the LCP arrays in external memory. This motivates our work here to design efficient checkers for big suffix and LCP arrays in external memory.

@@

Our contribution comprises two methods to probabilistically verify any given suffix and LCP arrays. In principle, Method A checks the lexical order and the LCP-value of two neighboring suffixes in the suffix array by literally comparing the characters of their LCPs. For reducing the time complexity of a comparison between two sequences of characters, we use a Karp-Rabin fingerprinting technique to convert each sequence into a single integer, called fingerprint, and compare the fingerprints instead to check the equality of two sequences. The algorithm for Method A involves multiple scans and sorts on sets of $\mathcal{O}(n)$ fixed-size items. Its implementation in external memory suffers from a space bottleneck due to the large disk volume taken by each sort. 
To overcome this drawback, Method B first employs the fingerprinting technique to check a subset selected from the given suffix and LCP arrays, then it reuses the inducing process of the IS-based construction algorithm SA-IS \cite{Nong11} to produce the final suffix and LCP arrays from the verified subset and literally compares them with the input arrays to ensure the correctness of the latter. Our experiments indicate that the peak disk use of the programs for Algorithm~\ref{alg:2} designed by Method B is only about half as that of the program for Algorithm~\ref{alg:1} designed by Method A.

The remainder of this paper is organized as follows. Sections~\ref{sec:method1} and~\ref{sec:method2} describe the two methods and their algorithmic designs. Section~\ref{sec:experiment} conducts an experimental study for performance evaluation of our programs for these algorithms. Section~\ref{sec:conclusion} gives some concluding remarks.

@@

Given a string $x[0, n-1]$ drawn from a constant or integer alphabet $\Sigma$ of size $O(1)$ or $O(n)$, respectively, the suffix array of $x$, denoted by $sa$, is a permutation of $\{0, 1, ..., n - 1\}$ such that ${\sf suf}(sa[i]) < {\sf suf}(sa[j])$ for $i, j \in [0, n)$ and $i < j$, where ${\sf suf}(sa[i])$ and ${\sf suf}(sa[j])$ are two suffixes starting with $x[sa[i]]$ and $x[sa[j]]$, respectively. Particularly, we say ${\sf suf}(sa[j])$ is a lexical neighbor of ${\sf suf}(sa[i])$ if $|i - j| = 1$. The LCP array of $x$, denoted by $lcp$, consists of $n$ integers, where $lcp[0]=0$ and $lcp[i]$ records the LCP-value of ${\sf suf}(sa[i])$ and ${\sf suf}(sa[i - 1])$ for $i \in [1, n)$.

@@

According to the above definitions, we give in Lemma~\ref{lemma:1} the sufficient and necessary conditions for checking suffix and LCP arrays. Notice that the lexical order and the LCP-value of any two suffixes in $x$ can be computed by literally comparing their characters rightward. For convenience, we append a virtual character to $x$ and assume it to be lexicographically smaller than any characters in $\Sigma$. Because all the suffixes differ in length and end with the virtual character, any two suffixes are different.

@@

If we directly compare the characters of two suffixes, the worst case time is $O(n)$. An alternative is to exploit a perfect hash function to convert each substring into a single integer such that any two substrings have a common hash value if and only if they are literally equal to each other, hence we can compare the hash values of two substrings instead to check their equality. Taking into account the high difficulty of finding such a perfect hash function to meet this requirement, we prefer using a Karp-Rabin fingerprinting function~\cite{Karp1987} to transform a substring into an integer called fingerprint. To be specific, suppose $L$ is a prime and $\delta$ is a number randomly chosen from $[1, L)$, the fingerprint ${\sf fp}(i, j)$ for a substring $x[i, j]$ can be iteratively calculated according to the formulas below as follows: scan $x$ rightward to iteratively compute ${\sf fp}(0, k)$ for all $k \in [0, n)$ using Formulas~\ref{formula:1}-\ref{formula:2}, record ${\sf fp}(0, i - 1)$ and ${\sf fp}(0, j)$ during the calculation and subtract the former from the latter to obtain ${\sf fp}(i, j)$ using Formula~\ref{formula:3}.

@@

Notice that two equal substrings always share a common fingerprint, but the inverse is not true. It has been proved in~\cite{Karp1987} that the probability of a false match can be reduced to a negligible level by setting $L$ to a large value\footnote{This property is utilized in~\cite{Bille2013} to design a probabilistic algorithm for computing a sparse suffix array.}. Hence, we have:

@@

Fig.~\ref{fig:example} gives an illustrating example for utilizing Corollary~\ref{corollary:1} to check the input suffix and LCP arrays. Given that $L = 197$ and $\delta = 101$, lines 4-8 compute ${\sf fp}(0, p)$ iteratively according to Formulas~\ref{formula:1}-\ref{formula:2}. Then, lines 10-16 use these values to compute the fingerprints for all the target substrings. In more detail, consider the leftmost pair of neighboring suffixes in $sa$, that is ${\sf suf}(sa[0])$ and ${\sf suf}(sa[1])$, the substrings indicated by their LCP-value are $x[sa[0], sa[0] + lcp[1] - 1]$ and $x[sa[1], sa[1] + lcp[1] - 1]$, respectively. According to Formula~\ref{formula:3}, ${\sf fp}(sa[0], sa[0] + lcp[1] - 1)$ is equal to the difference between ${\sf fp}(0, sa[0] - 1)$ and ${\sf fp}(0, sa[0] + lcp[1] - 1)$, both of which have been calculated beforehand. Following the same way, ${\sf fp}(sa[1], sa[1] + lcp[1] - 1)$ is computed by reducing ${\sf fp}(0, sa[1] - 1)$ from ${\sf fp}(0, sa[1] + lcp[1] - 1)$. Hence, we obtain the fingerprints for these two substrings in lines 10-12 and see that they are equal to each other.

@@

	\includegraphics[width = 0.9\columnwidth]{example}
	\caption{An example for computing and comparing fingerprints for substrings specifed by the LCP-Values of neighboring suffixes in the suffix array. \label{fig:example}}	
\end{figure}

@@

It is clear that the above algorithm consumes $\mathcal{O}(n)$ time and space when implemented in internal memory. However, if the two auxiliary arrays cannot be wholly accommodated into RAM during the execution of S2, it suffers from a performance degradation caused by frequent random accesses to disks. Assume that $x$, $sa$ and $lcp$ are stored in external memory, we design Algorithm~\ref{alg:1} for conducting these I/O operations in a disk-friendly way. The main idea is to first sort data in their access order and then visit them by sequential reads. For the purpose, Algorithm~\ref{alg:1} first scans $sa$ and $lcp$ to produce $ST_1, ST_2, ST_3$ and sorts their tuples by 1st component in ascending order at the very beginning (lines~\ref{alg:1:a}-\ref{alg:1:b}). Afterward, it iteratively computes the fingerprints of all the prefixes according to Formulas~\ref{formula:1}-\ref{formula:2} and assigns them to the sorted tuples as following (lines~\ref{alg:1:c}-\ref{alg:1:d}): when figuring out ${\sf fp}(0, i - 1)$, extract each tuple $e$ with $e.1st = i$ from $ST_1/ST_2/ST_3$, update $e$  with ${\sf fp}(0, i - 1)$, and then forward $e$ to $ST_1'/ST_2'/ST_3'$. Because the 1st components of the tuples in $ST_1$ constitute a copy of $sa$, the algorithm checks condition (1) when scanning these tuples in their sorted order. Finally, it sorts the updated tuples back to their original order~(line~\ref{alg:1:e}) and visits them  sequentially to check conditions (2)-(3) following the same way of S2~(lines~\ref{alg:1:f}-\ref{alg:1:g}).

@@

			$fp := fp \cdot \delta + x[i] \! \mod \! P$ \hspace{5cm} // $x[n]$ is the virtual character
		} \label{alg:1:d}

@@

			\If{$\hat{fp_1} \ne \hat{fp_2}$ {\sf or} $ch_1 \ge ch_2$}{

@@

The last point to be mentioned here is how to obtain the value of $\delta^{lcp[i]}$ quickly when computing $\hat{fp_1}$ and $\hat{fp_2}$ in lines~\ref{alg:1:j} and~\ref{alg:1:k}. One method is to keep a lookup table in internal memory to store $\delta^{lcp[i]}$ for all $i \in [0, n)$. This can answer the question in constant time, but it is impractical if the table is larger than the available memory size. Notice that the LCP of any two suffixes is smaller than $n$, thus we can return the answer in $\mathcal{O}(\lceil \log_2{n} \rceil)$ time using $\mathcal{O}(\lceil \log_2{n} \rceil)$ internal memory space based on the following idea. Suppose $e$ is an integer from $[0, n)$, it can be broken down into the form of $\Sigma_{i = 0}^{\lceil \log_2{n} \rceil}{k_i \cdot 2^i}$ by performing at most $\lceil \log_2{n} \rceil$ divisions, where $k_0, k_1, ..., k_{\lceil \log_2{n} \rceil}$ are integers from $\{0, 1\}$ . Thereby, we have $\delta^e = \Pi_{i = 0}^{\lceil \log_2{n} \rceil}{\delta}^{k_i \cdot 2^i}$ by replacing $e$ with its decomposition, where the expression on the right side of the equation can be easily computed with $\{{\delta}^{1}, {\delta}^{2}, \dots, {\delta}^{2^{\lceil \log_2{n} \rceil}} \}$ already known.

@@

Algorithm~\ref{alg:1} performs multiple scans and sorts on the arrays of $\mathcal{O}(n)$ fixed-size tuples residing on disks. Given an external memory model with RAM size $M$, disk size $D$ and block size $B$, all are in words, the time and I/O complexities for each scan are $\mathcal{O}(n)$ and $\mathcal{O}(n / B)$, respectively, while those for each sort are $\mathcal{O}(n\log_{M/ B}(n / B))$ and $\mathcal{O}((n / B)\log_{M / B}(n / B))$, respectively~\cite{Arge2013}. Algorithm~\ref{alg:1} reaches its peak disk use when sorting tuples in lines~\ref{alg:1:b} and~\ref{alg:1:e}. Suppose the input string and the suffix/LCP array are encoded using $\alpha$- and $\beta$-byte integers, respectively, and each fingerprint is represented by a $\gamma$-byte integer, it takes $(2 \cdot (2 \cdot \alpha + \beta)) \cdot n$ space for sorting $ST_1/ST_1'$ and $(2 \cdot (2 \cdot \alpha + \beta + \gamma)) \cdot n$ for sorting $ST_2/ST_2'$ and $ST_3/ST_3'$. To save disk space, our program implementing the algorithm sorts $ST_1/ST_1'$, $ST_2/ST_2'$ and $ST_3/ST_3'$ separately and performs a single scan over $x$ for each of them to obtain the target fingerprints, resulting in smaller space requirement but larger running time. The experimental study in Section~\ref{sec:experiment} indicates that this program is still space consuming, its peak disk use is 40 bytes per input character.

@@

In this part, we describe an alternative checking method based on the induced sorting principle. Compared with Algorithm~\ref{alg:1}, our programs for Algorithm~\ref{alg:2} designed by this method take around half disk space on real-world datasets. Before the presentation, we introduce some symbols and notations used in the following paragraphs.

@@

{\em Suffix and LCP buckets.} Suppose $sa$ is correct, all the suffixes in $sa$ are naturally partitioned into multiple buckets and those of a common heading character are grouped into a single bucket that occupies a contiguous interval in $sa$. Each bucket can be further divided into two sub-buckets, where the left and the right parts only contain L- and S-type suffixes, respectively. For short, we use ${\sf sa\_bkt}(c)$ to denote the bucket storing suffixes starting with character $c$ and ${\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$ to denote its left/right sub-bucket. Accordingly, $lcp$ can be also split into multiple buckets, where ${\sf lcp\_bkt}(c)/{\sf lcp\_bkt_L}(c)/{\sf lcp\_bkt_S}(c)$ stores the LCP-values of suffixes in ${\sf sa\_bkt}(c)/{\sf sa\_bkt_L}(c)/{\sf sa\_bkt_S}(c)$.

{\em Suffix and LCP arrays for S*-type suffixes.} Given that the number of S*-type suffixes is $n_1$, $sa^*[0, n_1)$ stores all the S*-type suffixes arranged in lexical order, while $lcp^*[0] = 0$ and $lcp^*[i]$ records the LCP-value of ${\sf suf}(sa^*[i])$ and ${\sf suf}(sa^*[i - 1])$ for $i \in [1, n_1)$.

@@

The induced sorting principle has been extensively used to design efficient algorithms for constructing the suffix and LCP arrays on internal and external memory models. An IS-based construction algorithm mainly consists of a reduction phase for computing $sa^*$ and $lcp^*$, followed by an induction phase for inducing $sa$ and $lcp$ from $sa^*$ and $lcp^*$~\footnote{In the interest of completeness, we give an overview of the induction phase in Appendix~\ref{sec:appendix}.}. Given that $sa^*$ and $lcp^*$ are already known, we can produce the final suffix and LCP arrays by calling the inducing process of an existing IS-based construction algorithm. This enlightens us to design a checker based on Lemma~\ref{lemma:2}.

@@

\begin{enumerate}[(1)]
	\item
	$sa^*$ and $lcp^*$ are both correct.
	\item
	$sa = sa'$ and $lcp = lcp'$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by the IS method.
\end{enumerate}
\end{Lemma}

Similar to Method A, the fingerprinting technique can be employed to probabilistically check the correctness of $sa^*$ and $lcp^*$. An external-memory algorithm for checking the conditions in Corollary~\ref{corollary:2} is given in the next subsection.

\begin{Corollary} \label{corollary:2}
Both $sa[0, n)$ and $lcp[0, n)$ are correct with a high probability given the following conditions, for all $i \in [1, n_1)$ and $j \in [0, n)$:
\begin{enumerate}[(1)]
	\item
	$x[sa^*[i]]$ is S*-type, and $sa^*[i] \ne sa^*[k]$ for all $k \in [0, n_1)$ and $k \ne i$.
	\item
	${\sf fp}(sa^*[i], sa^*[i] + lcp^*[i] - 1) = {\sf fp}(sa^*[i - 1], sa^*[i - 1] + lcp^*[i] - 1)$.
	\item
	$x[sa^*[i] + lcp^*[i]] > x[sa^*[i - 1] + lcp^*[i]]$.
	\item
	$sa[j] = sa'[j]$ and $lcp[j] = lcp'[j]$ for $j \in [0, n)$, where $sa'$ and $lcp'$ are induced from $sa^*$ and $lcp^*$ by the IS method.
\end{enumerate}
\end{Corollary}

@@

	sort tuples in $ST_2$ by the 2nd component \label{alg:2:g}

@@

	\If{${\sf CheckByFP}(x, sa^*, lcp^*, n_1) = false$}{  \label{alg:2:i}

		\Return false \hspace{1cm} // conditions (2) or (3) is violated
	}\label{alg:2:j}

@@

			\Return false \hspace{1cm}	// condition (4) is violated
		}	
	}\label{alg:2:m}
